Anony\emph{mice}d shareable data: Using \emph{mice} to create and analyze multiply imputed synthetic data sets

### 2021-11-02 manuscript/Manuscript.Rmd
Add final revisions

If the goal is to create synthetic versions of a data set that suffers from missingness, the imputation procedure becomes more complex. A two-step imputation procedure has been proposed by @reiter2004simultaneous. In the first step, the missing data is imputed according to the principles of multiple imputation for missing data, while in a second step, the synthetic data is imputed conform the corresponding procedure. Ideally, these steps could be combined using `mice`, conditioning the imputations on both the missingness and the values that are to be synthesized. In practice, replacing observed and missing cells in a data set by synthetic values into the object `syn`, given the same imputation parameters as before, can be realized by the following code execution.

```{r eval = FALSE}
syn <- mice(data.with.missings.to.synthesize, 
            m = 10,
            maxit = 7, 
            where = matrix(TRUE, 
                           nrow = nrow(data.with.missings.to.synthesize),
                           ncol = ncol(data.with.missings.to.synthesize)))
```

@@

$$
\bar{u}_m = \sum_{l = 1}^m \frac{u^{(l)}}{m},
$$
where $u^{(l)}$ denotes the variance estimate in the $l$th synthetic data set. These pooling rules are implemented in `mice` as the function `pool.syn`. @reiter_kinney_2012 proved that synthesizing data from the posterior predictive distribution is not required for these pooling rules to hold. In fact, the parameters of an imputation model can be fixed to their maximum likelihood estimates or their posterior modes, and synthetic values can be drawn after plugging these values into the imputation models. This observation as been taken a step further by @raab_practical_2016, who developed pooling rules that do not require multiple synthetic data sets and performed highly similar to the aforementioned pooling rules. However, because these pooling rules can only be used when the complete data (rather than a subset of the data) is synthesized, we adhere to the originally proposed pooling rules by @reiter_partially_inference_2003, who defined the total variance in terms of sampling variance and between-imputation variance.

@@

@@

library(xtable)

bind_rows(Boostrapped = boot_stats_cont, 
          Synthetic = syn_stats_cont, 
          .id = "Class") %>%
  ggplot(aes(x = mean, col = Class, fill = Class)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~name, scales = "free") +
  theme_classic() +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Mean",
       y = "Density",
       title = "")
```

bind_rows(Boostrapped = boot_stats_cat, 
          Synthetic = syn_stats_cat, 
          .id = "Class") %>%
  ggplot(aes(y = value, x = `Number of observations`, fill = Class)) +
  geom_boxplot(position = "dodge", width = 0.8) +
  facet_wrap(~name, scales = "free") +
  theme_classic() +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "Number of observations per category",
       y = "Category",
       title = "")
```

```{r, include = FALSE}
library(xtable)
bind_rows(true, setNames(synth, names(true))) %>% 
           arrange(vars) %>% .[, c(2:5, 8:9, 11:12)] %>% round(3) %>% 
  xtable
```
<!-- % latex table generated in R 4.1.0 by xtable 1.8-4 package -->
<!-- % Mon Oct  4 16:47:59 2021 -->
\begin{table}[ht]
\caption{Univariate descriptives for the true data and $m=5$ pooled univariate descriptives for the synthetic data over 1000 simulations. Variable names followed by an asterix ($^*$) are categorical.}
\centering
\begin{tabular}{rrrrrrrrr}
  \hline
 & n & mean & sd & median & min & max & skew & kurtosis \\ 
  \hline
original age & 748 & 9.16 & 6.89 & 10.50 & 0.04 & 21.18 & -0.03 & -1.56 \\ 
  synthetic age & 748 & 9.15 & 6.89 & 10.49 & 0.04 & 20.96 & -0.03 & -1.55 \\ 
  original hgt & 748 & 131.10 & 46.52 & 145.75 & 50.00 & 198.00 & -0.30 & -1.47 \\ 
  synthetic hgt & 748 & 131.06 & 46.50 & 145.32 & 50.69 & 197.16 & -0.30 & -1.47 \\ 
  original wgt & 748 & 37.12 & 26.03 & 34.55 & 3.14 & 117.40 & 0.38 & -1.03 \\ 
  synthetic wgt & 748 & 37.09 & 26.00 & 34.44 & 3.35 & 112.26 & 0.38 & -1.03 \\ 
  original bmi & 748 & 18.04 & 3.04 & 17.45 & 11.73 & 31.74 & 1.14 & 1.79 \\ 
  synthetic bmi & 748 & 18.05 & 3.08 & 17.48 & 11.49 & 32.37 & 1.11 & 1.85 \\ 
  original hc & 748 & 51.62 & 5.86 & 53.10 & 33.70 & 65.00 & -0.91 & 0.12 \\ 
  synthetic hc & 748 & 51.61 & 5.86 & 53.18 & 34.38 & 62.85 & -0.91 & 0.12 \\ 
  original gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.36 \\ 
  synthetic gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.35 \\ 
  original phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.25 \\ 
  synthetic phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.24 \\ 
  original tv & 748 & 8.43 & 8.12 & 3.00 & 1.00 & 25.00 & 0.85 & -0.78 \\ 
  synthetic tv & 748 & 8.42 & 8.11 & 3.19 & 1.00 & 25.00 & 0.85 & -0.77 \\ 
  original reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.77 \\ 
  synthetic reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.76 \\ 
   \hline
\end{tabular}
\end{table}

>>>>

If the goal is to create synthetic versions of a data set that suffers from missingness, the imputation procedure becomes more complex. A two-step imputation procedure has been proposed by @reiter2004simultaneous. In the first step, the missing data is imputed according to the principles of multiple imputation for missing data, while in a second step, the synthetic data is imputed conform the corresponding procedure. Ideally, these steps could be combined using `mice`, conditioning the imputations on both the missingness and the values that are to be synthesized. In practice, replacing observed and missing cells in a data set by synthetic values into the object `syn.mis`, given the same imputation parameters as before, can be realized by the following code execution.

```{r eval = FALSE}
syn.mis <- mice(data.with.missings.to.synthesize, 
                m = 10,
                maxit = 7, 
                where = matrix(TRUE, 
                               nrow = nrow(data.with.missings.to.synthesize),
                               ncol = ncol(data.with.missings.to.synthesize)))
```

@@

$$
\bar{u}_m = \sum_{l = 1}^m \frac{u^{(l)}}{m},
$$
where $u^{(l)}$ denotes the variance estimate in the $l$th synthetic data set. These pooling rules are implemented in `mice` as the function `pool.syn()`. Returning to our previous example, pooling the parameter estimates of a linear model in which the dependent variable `DV` is regressed on predictors `IV1` and `IV2` fitted on each of the synthetic data sets is as straightforward as the following code block.

```{r, eval = FALSE}
fits <- with(syn, lm(DV ~ IV1 + IV2))
pool.syn(fits)
```

@reiter_kinney_2012 proved that synthesizing data from the posterior predictive distribution is not required for these pooling rules to hold. In fact, the parameters of an imputation model can be fixed to their maximum likelihood estimates or their posterior modes, and synthetic values can be drawn after plugging these values into the imputation models. This observation as been taken a step further by @raab_practical_2016, who developed pooling rules that do not require multiple synthetic data sets and performed highly similar to the aforementioned pooling rules. However, because these pooling rules can only be used when the complete data (rather than a subset of the data) is synthesized, we adhere to the originally proposed pooling rules by @reiter_partially_inference_2003, who defined the total variance in terms of sampling variance and between-imputation variance.

@@

```{r univ-stats, include = FALSE, cache = TRUE}
library(xtable)

bind_rows(true, setNames(synth, names(true))) %>% 
           arrange(vars) %>% .[, c(2:5, 8:9, 11:12)] %>% round(3) %>% 
  xtable
```

<!-- % latex table generated in R 4.1.0 by xtable 1.8-4 package -->
<!-- % Mon Oct  4 16:47:59 2021 -->
\begin{table}[ht]
\caption{Univariate descriptives for the true data and $m=5$ pooled univariate descriptives for the synthetic data over 1000 simulations. Variable names followed by an asterix ($^*$) are categorical.}
\centering
\begin{tabular}{rrrrrrrrr}
  \hline
 & n & mean & sd & median & min & max & skew & kurtosis \\ 
  \hline
original age & 748 & 9.16 & 6.89 & 10.50 & 0.04 & 21.18 & -0.03 & -1.56 \\ 
  synthetic age & 748 & 9.15 & 6.89 & 10.49 & 0.04 & 20.96 & -0.03 & -1.55 \\ 
  original hgt & 748 & 131.10 & 46.52 & 145.75 & 50.00 & 198.00 & -0.30 & -1.47 \\ 
  synthetic hgt & 748 & 131.06 & 46.50 & 145.32 & 50.69 & 197.16 & -0.30 & -1.47 \\ 
  original wgt & 748 & 37.12 & 26.03 & 34.55 & 3.14 & 117.40 & 0.38 & -1.03 \\ 
  synthetic wgt & 748 & 37.09 & 26.00 & 34.44 & 3.35 & 112.26 & 0.38 & -1.03 \\ 
  original bmi & 748 & 18.04 & 3.04 & 17.45 & 11.73 & 31.74 & 1.14 & 1.79 \\ 
  synthetic bmi & 748 & 18.05 & 3.08 & 17.48 & 11.49 & 32.37 & 1.11 & 1.85 \\ 
  original hc & 748 & 51.62 & 5.86 & 53.10 & 33.70 & 65.00 & -0.91 & 0.12 \\ 
  synthetic hc & 748 & 51.61 & 5.86 & 53.18 & 34.38 & 62.85 & -0.91 & 0.12 \\ 
  original gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.36 \\ 
  synthetic gen* & 748 & 2.53 & 1.59 & 2.00 & 1.00 & 5.00 & 0.52 & -1.35 \\ 
  original phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.25 \\ 
  synthetic phb* & 748 & 2.75 & 1.86 & 2.00 & 1.00 & 6.00 & 0.56 & -1.24 \\ 
  original tv & 748 & 8.43 & 8.12 & 3.00 & 1.00 & 25.00 & 0.85 & -0.78 \\ 
  synthetic tv & 748 & 8.42 & 8.11 & 3.19 & 1.00 & 25.00 & 0.85 & -0.77 \\ 
  original reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.77 \\ 
  synthetic reg* & 748 & 3.02 & 1.14 & 3.00 & 1.00 & 5.00 & -0.08 & -0.76 \\ 
   \hline
\end{tabular}
\end{table}

bind_rows(Bootstrapped = boot_stats_cont, 
          Synthetic = syn_stats_cont, 
          .id = "Class") %>%
  ggplot(aes(x = mean, col = Class, fill = Class)) +
  geom_density(alpha = 0) +
  facet_wrap(~name, scales = "free") +
  theme_classic() +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  scale_color_brewer(palette = "Set1", direction = -1) +
  labs(x = "Mean",
       y = "Density",
       title = "")
```

bind_rows(Bootstrapped = boot_stats_cat, 
          Synthetic = syn_stats_cat, 
          .id = "Class") %>%
  ggplot(aes(y = value, x = `Number of observations`, fill = Class)) +
  geom_boxplot(position = "dodge", width = 0.8) +
  facet_wrap(~name, scales = "free") +
  theme_classic() +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Set1", direction = -1) +
  labs(x = "Number of observations per category",
       y = "Category",
       title = "")
```

@@

### 2021-11-02 manuscript/Manuscript.tex
Add final revisions

If the goal is to create synthetic versions of a data set that suffers
from missingness, the imputation procedure becomes more complex. A
two-step imputation procedure has been proposed by
\citet{reiter2004simultaneous}. In the first step, the missing data is
imputed according to the principles of multiple imputation for missing
data, while in a second step, the synthetic data is imputed conform the
corresponding procedure. Ideally, these steps could be combined using
\texttt{mice}, conditioning the imputations on both the missingness and
the values that are to be synthesized. In practice, replacing observed
and missing cells in a data set by synthetic values into the object
\texttt{syn}, given the same imputation parameters as before, can be
realized by the following code execution.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{syn }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(data.with.missings.to.synthesize, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{7}\NormalTok{, }
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(data.with.missings.to.synthesize),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(data.with.missings.to.synthesize)))}
\end{Highlighting}
\end{Shaded}

@@

\[
\bar{u}_m = \sum_{l = 1}^m \frac{u^{(l)}}{m},
\] where \(u^{(l)}\) denotes the variance estimate in the \(l\)th
synthetic data set. These pooling rules are implemented in \texttt{mice}
as the function \texttt{pool.syn}. \citet{reiter_kinney_2012} proved
that synthesizing data from the posterior predictive distribution is not
required for these pooling rules to hold. In fact, the parameters of an
imputation model can be fixed to their maximum likelihood estimates or
their posterior modes, and synthetic values can be drawn after plugging
these values into the imputation models. This observation as been taken
a step further by \citet{raab_practical_2016}, who developed pooling
rules that do not require multiple synthetic data sets and performed
highly similar to the aforementioned pooling rules. However, because
these pooling rules can only be used when the complete data (rather than
a subset of the data) is synthesized, we adhere to the originally
proposed pooling rules by \citet{reiter_partially_inference_2003}, who
defined the total variance in terms of sampling variance and
between-imputation variance.

>>>>

If the goal is to create synthetic versions of a data set that suffers
from missingness, the imputation procedure becomes more complex. A
two-step imputation procedure has been proposed by
\citet{reiter2004simultaneous}. In the first step, the missing data is
imputed according to the principles of multiple imputation for missing
data, while in a second step, the synthetic data is imputed conform the
corresponding procedure. Ideally, these steps could be combined using
\texttt{mice}, conditioning the imputations on both the missingness and
the values that are to be synthesized. In practice, replacing observed
and missing cells in a data set by synthetic values into the object
\texttt{syn.mis}, given the same imputation parameters as before, can be
realized by the following code execution.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{syn.mis }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(data.with.missings.to.synthesize, }
                \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
                \AttributeTok{maxit =} \DecValTok{7}\NormalTok{, }
                \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }
                               \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(data.with.missings.to.synthesize),}
                               \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(data.with.missings.to.synthesize)))}
\end{Highlighting}
\end{Shaded}

@@

\[
\bar{u}_m = \sum_{l = 1}^m \frac{u^{(l)}}{m},
\] where \(u^{(l)}\) denotes the variance estimate in the \(l\)th
synthetic data set. These pooling rules are implemented in \texttt{mice}
as the function \texttt{pool.syn()}. Returning to our previous example,
pooling the parameter estimates of a linear model in which the dependent
variable \texttt{DV} is regressed on predictors \texttt{IV1} and
\texttt{IV2} fitted on each of the synthetic data sets is as
straightforward as the following code block.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{fits }\OtherTok{\textless{}{-}} \FunctionTok{with}\NormalTok{(syn, }\FunctionTok{lm}\NormalTok{(DV }\SpecialCharTok{\textasciitilde{}}\NormalTok{ IV1 }\SpecialCharTok{+}\NormalTok{ IV2))}
\FunctionTok{pool.syn}\NormalTok{(fits)}
\end{Highlighting}
\end{Shaded}

\citet{reiter_kinney_2012} proved that synthesizing data from the
posterior predictive distribution is not required for these pooling
rules to hold. In fact, the parameters of an imputation model can be
fixed to their maximum likelihood estimates or their posterior modes,
and synthetic values can be drawn after plugging these values into the
imputation models. This observation as been taken a step further by
\citet{raab_practical_2016}, who developed pooling rules that do not
require multiple synthetic data sets and performed highly similar to the
aforementioned pooling rules. However, because these pooling rules can
only be used when the complete data (rather than a subset of the data)
is synthesized, we adhere to the originally proposed pooling rules by
\citet{reiter_partially_inference_2003}, who defined the total variance
in terms of sampling variance and between-imputation variance.
